Domain-specific working memory loads selectively increase negative interpertations of surprised facial expressions

### 2019-09-24 Manuscript/Manuscript.Rmd
Updated intro

### load necessary libraries ###
suppressPackageStartupMessages(library(readxl)) 
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(openxlsx))
suppressPackageStartupMessages(library(Rmisc))
suppressPackageStartupMessages(library(broom))
```
# Introduction
Completion of even simple tasks in daily life require regular recruitment of working memory, and this often happens without consideration of how these working memory demands may affect concurrent processes. Many types of information can be held in working memory: strings of numbers, neutral and emotional images, stimulus characteristics like brightness, spatial locations, and more. There is debate in the literature around the neurophysiological substrates of working memory and whether this construct is a unified system or comprised of several domain-specific subsystems (see Baddeley, 1986; Shah & Mikaye, 1996). Recent work suggests though that working memory does indeed contain domain-specific subsystems (e.g., affective working memory; see @mikels_affective_2019)--or at least that interactions between working memory and visual or affective tasks are variable depending on working memory content. Similar domain-specific effects have been observed in a number of other lines of research, including those outside of the working memory domain. For example, the Stroop task [@stroop_studies_1935] has been modified to include emotional stimuli [@whalen_emotional_2006] which has pronounced effects when the emotional words are population specific (e.g., trauma words in a PTSD sample). Further, neuroimaging work supports the notion that separate systems handle attentional biasing for domain-specific (emotional vs. non-emotional) task relevancy [@egner_dissociable_2008]. The present work aims to clarify the interaction of working memory demands on a concurrent and ecologically valid emotionoal task (i.e., interpretations of emotional facial expressions), specifically when that content varies in both amount and affective quality.

There is a large body of research concerning the effects of distractor and task irrelevant stimuli on performance in a variety of tasks [cite, cite]. Notably, domain-specific interference may further exacerbate these effects compared to domain-general stimuli [@gruber_effects_2001]. Within an affect-specific context, the literature reports conflicting results for the role of affective stimuli on working memory processing (see @jackson_enhanced_2009 and @mather_emotional_2006), with some suggesting ehnaced recall for emotional stimuli and others reporting impairment in working memory processes. The relationship between these systems is bidirectional though, and often times working memory affects affective processing. One such way is through dampening emotional responses, and converging work suggests that cognitive demands can mitigate negative emotional effects [@dillen_tuning_2008]. One explanation for mitigated emotional responding during concurrent working memory demands is load theory. This theory posits that cognitive resources are finite, thus limiting the availability for simultaneous processing under large task demands [@lavie_load_2004]. One limitation to this previous work is that the affective processes under study involved clearly interpretable stimuli/feelings. A more malleable affective process, such as interpreting ambiguously valenced emotional facial expressions, may offer better understanding of how concurrent working memory demands affect emotional processes.

Facial expressions are rich with affective information, and correctly interpreting these social cues is critical for successfully navigating the social world. Often, facial expressions serve as a clear social signal, but this is not always the case. While a smile likely expresses a positive affective state, other cues are not so clear. For instance, a surprised expression could signal either a positive (e.g., winning the lottery) or negative (e.g., seeing a snake in the woods) affective state in the expresser. When contextual information is limited, individuals differ in their tendency to interpret surprised facial expressions as positive or negative. Importantly, this affective bias extends beyond facial expressions, as individuals often show a similar bias to surprised faces as they do for ambiguous scenes [@neta_neural_2013] or even words (Harp, Petro, Brown & Neta, in prep). This bias towards positive or negative interpretations is known as one's valence bias and myriad factors contribute to one's bias.

@@

In the present study, we aim to test the effects of low and high working memory loads in both emotional and neutral domains. We expect that trials in which participants are maintaining an emotional working memory load will be more negative than neutral trials. Further, we predict that higher working memory laod trials, specifically in the emotional domain, will result in even more exaggerated negative interpretations.

@@

>>>>

### load necessary libraries ###
suppressPackageStartupMessages(library(readxl)) 
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(openxlsx))
suppressPackageStartupMessages(library(Rmisc))
suppressPackageStartupMessages(library(broom))
```
# Introduction
## Working memory and load theory
Despite extensive research on the interaction of working memory and affective processes, there is still debate in the literature concerning how this cognitive process and emotional processes affect one another. Directly comparing working memory and self-regulation of emotional responses, Schmeichel and colleagues -[@schmeichel_working_2008] reported that individuals with higher levels of working memory capacity demonstrated improved self-regulation towards the emotional stimuli. This suggests a connection--perhaps through some shared resource pool--between mitigated emotional responding and larger working memory resource availablility. Other work has focused on the effects of moods or affective states on working memory performance. For instance, some reports claim that both positive and negative mood interfere with working memory (Eyesenck and Calvo, 1992); however, others suggest benefits of positive mood on working memory [@yang_positive_2012]. Similarly, working memory processes may alter concurrent affective processes. For instance, actively engaging working memory can mitigate emotional responses. Recent neuroimaging work reports that negative emotional responses decrease as the cognitive demands of a working memory task increase [@dillen_tuning_2008]. Additionally, depending on one's levels of trait-rumination, low trait-ruminators are better able to return heart rate and blood pressure to baseline levels when provided with a distractor task following an anger induction [@gerin_role_2006]. Together, these studies suggest a resource competition between cognitive and emotional processing; that is, when cognitive load demands are high (e.g., during active working memory maintenance), there are fewer resources available for other (i.e., affective) processes.

While previous work clearly suggests an inhibitory role for cognitive demands on emotional responses, researchers have primarily focused on emotional responses to clearly valenced emotional stimuli. For instance, Schmeichel and colleagues -[@schmeichel_working_2008] showed participants videos intended to elicit strong negative responses (e.g., disgust) or positive responses (e.g., humor), while others have focused on comparing responses to neutral and negative sitmuli [@dillen_tuning_2008], or even reducing symptoms of disorders with affective symptoms, such as anxiety [@spangler_little_2017]. However, many emotional appraisals in day-to-day life are more nuanced than those invoked by many of the types of negative stimuli one might encounter in the lab (e.g., snakes, mutilated bodies). For example, one may appraise the content of a billboard displaying a large order of french fries as either negative or positive depending on whether or not consuming that food would be (in)congruent with one's current goals. This emotional appraisal is completed under concurrent laod demands--that is, the perceiver must process both the emotional stimulus (i.e., the fries) as well as actively maintain their current goal state. If maintaining a goal state or some stimulus in working memory results in a large cognitive (not perceptual) load, then  Lavie's -[@lavie_load_2004] load theory posits that less executive resources are available to regulate incoming perceptual information. In turn, this could mean that more readily processed perceptual input (e.g., low spatial frequency images; LSF) might drive decisions under working memory loads.

Emotional stimuli readily capture attention compared to neutral stimuli, even in participants with amygdala damage [@hodsol_attentional_2011; @piech_attentional_2011]. Given emotional stimuli's priority position in the information processing stream, it may be that cognitive loads with emotional content, compared to neutral, differentially affect concurrent emotional appraisals. Indeed, negative emotional content slows performance on an n-back task [@kensinger_effect_2003]. Other domain-specific effects have been observed in many lines of research, including those beyond the working memory domain. For example, the Stroop task [@stroop_studies_1935], a common measurement tool for inhibitory control, has been modified to include both emotional and non-emotional (neutral) stimuli [@whalen_emotional_2006] which has pronounced effects when the emotional words are population specific (e.g., trauma words in a PTSD sample). Neuroimaging work also supports the notion that separate systems handle attentional biasing for domain-specific (emotional vs. non-emotional) task relevancy [@egner_dissociable_2008]. Given the behavioral and neurological evidence for dissociations of emotional and non-emotional information domains in executive functions (e.g., working memory, inhibitory control), the present work aims to clarify the interaction of emotional and non-emotional visual working memory demands on concurrent emotional judgments.

<!-- Frequently, tasks in daily life require recruitment of working memory, but active working memory demands may affect concurrent processing. Many types of information can be held in working memory: strings of numbers, images, stimulus characteristics like brightness, spatial locations, and more. There is debate in the literature around the substrates of working memory and how this seemingly unitary construct is distributed across domain-specific subsystems (see [@baddeley_working_1986; @shah_separability_1996]). Recent work suggests though that working memory may contain more domain-specific subsystems than previously believed (e.g., affective working memory; e.g., @mikels_affective_2019). Similar to the work of Shah and colleagues -[@baddeley_working_1986], Mikels (find acutal study) showed ?double dissociation? of the affective working memory system from a more visually based working memory (i.e., maintenance of image brightness in memory). Similar domain-specific effects have been observed in many other lines of research, including those beyond the working memory domain. For example, the Stroop task [@stroop_studies_1935], a common measurement tool for inhibitory control, has been modified to include both emotional and non-emotional (neutral) stimuli [@whalen_emotional_2006] which has pronounced effects when the emotional words are population specific (e.g., trauma words in a PTSD sample). Neuroimaging work also supports the notion that separate systems handle attentional biasing for domain-specific (emotional vs. non-emotional) task relevancy [@egner_dissociable_2008], suggesting a division of subsystems. Given the behavioral and neurological evidence for dissociations of information domains in working memory, task interference and more, the present work aims to clarify the interaction of emotional and non-emotional visual working memory demands on concurrent emotional judgments (i.e., interpretations of emotional facial expressions). -->

<!-- While it is well known that distractors can undermine task performance [@hagen_effect_1967; cite, cite), some evidence suggests a role for domain-specificity in task interference.  For instance, domain-specific interference further exacerbates these task effects compared to domain-general stimuli [@gruber_effects_2001]. Within an affect-specific context, the literature reports conflicting results for the role of affective stimuli on working memory processing (see @jackson_enhanced_2009 and @mather_emotional_2006), with some suggesting ehnaced recall for emotional stimuli and others reporting impairment in working memory processes. The relationship between these systems is bidirectional though, and often times working memory affects affective processing. One such way is through dampening emotional responses, and converging work suggests that cognitive demands can mitigate negative emotional effects [@dillen_tuning_2008]. One explanation for mitigated emotional responding during concurrent working memory demands is load theory. This theory posits that cognitive resources are finite, thus limiting the availability for simultaneous processing under large task demands [@lavie_load_2004]. One limitation to this previous work is that the affective processes under study involved clearly interpretable stimuli/feelings. A more malleable affective process, such as interpreting ambiguously valenced emotional facial expressions, may offer better understanding of how concurrent working memory demands affect emotional processes. -->

## The contribution of valence bias
<!-- Facial expressions are rich with affective information, and correctly interpreting these social cues is critical for successfully navigating the social world.--> As mentioned above, individuals differ in their tendency to interpret ambiguously valenced stimuli, like a tempting food item, as either positive or negative. <!-- Additionally, facial expressions can serve as ambiguously valenced stimuli. While most facial expressions send a clear social signal, this is not always the case. While a smile likely expresses a positive affective state, other cues are not so clear.--> This tendency is known as one's *valence bias*, and a growing body of work has used both facial expressions and scenes (or items) to quantify this individual difference (CITE). Biases towards negativity are related to psychopathology, such as depression and anxiety [@matthews_cognitive_2005]

<!--For instance, a surprised expression could signal either a positive (e.g., winning the lottery) or negative (e.g., seeing a snake in the woods) affective state in the expresser. When contextual information is limited, individuals differ in their tendency to interpret surprised facial expressions as positive or negative. Importantly, this affective bias extends beyond facial expressions, as individuals often show a similar bias to surprised faces as they do for ambiguous scenes [@neta_neural_2013] or even words (Harp, Petro, Brown & Neta, in prep). This bias towards positive or negative interpretations is known as one's valence bias and myriad factors contribute to one's bias. -->

@@

## The present study
In the present study, we aim to test the effects of low and high working memory loads in both emotional and neutral domains. We expect that trials in which participants are maintaining an emotional working memory load will be more negative than neutral trials. Further, we predict that higher working memory laod trials, specifically in the emotional domain, will result in even more exaggerated negative interpretations.

@@

Increased working memory demands (i.e., a higher cognitive load) do not always result in poorer performance on concurrent tasks. For instance Baddeley -[@baddeley_working_1986] reported that increasing load by adding digits to a rehearsed number did not affect accuracy on a concurrent verbal reasoning task--instead, there was an increase in the latency of response, a potential interference effect that did not alter overall accuracy.

### 2019-09-24 Manuscript/Manuscript.tex
Updated intro

Completion of even simple tasks in daily life require regular recruitment of working memory, and this often happens without consideration of how these working memory demands may affect concurrent processes. Many types of information can be held in working memory: strings of numbers, neutral and emotional images, stimulus characteristics like brightness, spatial locations, and more. There is debate in the literature around the neurophysiological substrates of working memory and whether this construct is a unified system or comprised of several domain-specific subsystems (see Baddeley, 1986; Shah \& Mikaye, 1996). Recent work suggests though that working memory does indeed contain domain-specific subsystems (e.g., affective working memory; see Mikels and Reuter-Lorenz (2019))--or at least that interactions between working memory and visual or affective tasks are variable depending on working memory content. Similar domain-specific effects have been observed in a number of other lines of research, including those outside of the working memory domain. For example, the Stroop task (Stroop, 1935) has been modified to include emotional stimuli (Whalen, Bush, Shin, \& Rauch, 2006) which has pronounced effects when the emotional words are population specific (e.g., trauma words in a PTSD sample). Further, neuroimaging work supports the notion that separate systems handle attentional biasing for domain-specific (emotional vs.~non-emotional) task relevancy (Egner, Etkin, Gale, \& Hirsch, 2008). The present work aims to clarify the interaction of working memory demands on a concurrent and ecologically valid emotionoal task (i.e., interpretations of emotional facial expressions), specifically when that content varies in both amount and affective quality.

There is a large body of research concerning the effects of distractor and task irrelevant stimuli on performance in a variety of tasks {[}cite, cite{]}. Notably, domain-specific interference may further exacerbate these effects compared to domain-general stimuli (Gruber, 2001). Within an affect-specific context, the literature reports conflicting results for the role of affective stimuli on working memory processing (see ({\textbf{???}}) and ({\textbf{???}})), with some suggesting ehnaced recall for emotional stimuli and others reporting impairment in working memory processes. The relationship between these systems is bidirectional though, and often times working memory affects affective processing. One such way is through dampening emotional responses, and converging work suggests that cognitive demands can mitigate negative emotional effects ({\textbf{???}}). One explanation for mitigated emotional responding during concurrent working memory demands is load theory. This theory posits that cognitive resources are finite, thus limiting the availability for simultaneous processing under large task demands (Lavie, Hirst, Fockert, \& Viding, 2004). One limitation to this previous work is that the affective processes under study involved clearly interpretable stimuli/feelings. A more malleable affective process, such as interpreting ambiguously valenced emotional facial expressions, may offer better understanding of how concurrent working memory demands affect emotional processes.

@@

@@

@@

\leavevmode\hypertarget{ref-gruber_effects_2001}{}%
Gruber, O. (2001). Effects of domain-specific interference on brain activation associated with verbal working memory task performance. \emph{Cerebral Cortex (New York, N.Y.: 1991)}, \emph{11}(11), 1047--1055. doi:\href{https://doi.org/10.1093/cercor/11.11.1047}{10.1093/cercor/11.11.1047}

@@

\leavevmode\hypertarget{ref-mikels_affective_2019}{}%
Mikels, J. A., \& Reuter-Lorenz, P. A. (2019). Affective working memory: An integrative psychological construct. \emph{Perspectives on Psychological Science: A Journal of the Association for Psychological Science}, \emph{14}(4), 543--559. doi:\href{https://doi.org/10.1177/1745691619837597}{10.1177/1745691619837597}

>>>>

\hypertarget{working-memory-and-load-theory}{%
\subsection{Working memory and load theory}\label{working-memory-and-load-theory}}

Despite extensive research on the interaction of working memory and affective processes, there is still debate in the literature concerning how this cognitive process and emotional processes affect one another. Directly comparing working memory and self-regulation of emotional responses, Schmeichel and colleagues -({\textbf{???}}) reported that individuals with higher levels of working memory capacity demonstrated improved self-regulation towards the emotional stimuli. This suggests a connection--perhaps through some shared resource pool--between mitigated emotional responding and larger working memory resource availablility. Other work has focused on the effects of moods or affective states on working memory performance. For instance, some reports claim that both positive and negative mood interfere with working memory (Eyesenck and Calvo, 1992); however, others suggest benefits of positive mood on working memory ({\textbf{???}}). Similarly, working memory processes may alter concurrent affective processes. For instance, actively engaging working memory can mitigate emotional responses. Recent neuroimaging work reports that negative emotional responses decrease as the cognitive demands of a working memory task increase ({\textbf{???}}). Additionally, depending on one's levels of trait-rumination, low trait-ruminators are better able to return heart rate and blood pressure to baseline levels when provided with a distractor task following an anger induction ({\textbf{???}}). Together, these studies suggest resource competition between cognitive tasks and emotional processing; that is, when cognitive load demands are high (i.e., during active working memory maintenance), there are fewer resources available for other (i.e., affective) processes.

While previous work clearly suggests an inhibitory role for cognitive demands on emotional responses, researchers have primarily focused on emotional responses to clearly valenced emotional stimuli. For instance, Schmeichel and colleagues -({\textbf{???}}) showed participants videos intended to elicit strong negative responses (e.g., disgust) or positive responses (e.g., humor), while others have focused on comparing responses to neutral and negative sitmuli ({\textbf{???}}), or even reducing symptoms of disorders with affective symptoms, such as anxiety ({\textbf{???}}). However, many emotional appraisals in day-to-day life are more nuanced than those invoked by many of the types of negative stimuli one might encounter in the lab (e.g., snakes, mutilated bodies). For example, one may appraise the content of a billboard displaying a large order of french fries as either negative or positive depending on whether or not consuming that food would be (in)congruent with one's current goals. This emotional appraisal is completed under concurrent laod demands--that is, the perceiver must process both the emotional stimulus (i.e., the fries) as well as actively maintain their current goal state. If maintaining a goal state or some stimulus in working memory results in a large cognitive (not perceptual) load, then Lavie's -(Lavie, Hirst, Fockert, \& Viding, 2004) load theory posits that less executive resources are available to regulate incoming perceptual information.

Emotional stimuli readily capture attention compared to neutral stimuli, even in participants with amygdala damage {[}({\textbf{???}}); piech\_attentional\_2011{]}. Given emotional stimuli's priority position in the processing stream, it may be that cognitive loads with emotional content, compared to neutral, differentially affect concurrent emotional appraisals. Indeed, negative emotional content slows performance on an n-back task ({\textbf{???}}). domain-specific effects have been observed in many other lines of research, including those beyond the working memory domain. For example, the Stroop task (Stroop, 1935), a common measurement tool for inhibitory control, has been modified to include both emotional and non-emotional (neutral) stimuli (Whalen, Bush, Shin, \& Rauch, 2006) which has pronounced effects when the emotional words are population specific (e.g., trauma words in a PTSD sample). Neuroimaging work also supports the notion that separate systems handle attentional biasing for domain-specific (emotional vs.~non-emotional) task relevancy (Egner, Etkin, Gale, \& Hirsch, 2008), suggesting a division of subsystems. Given the behavioral and neurological evidence for dissociations of information domains in working memory, task interference and more, the present work aims to clarify the interaction of emotional and non-emotional visual working memory demands on concurrent emotional judgments.

\hypertarget{the-contribution-of-valence-bias}{%
\subsection{The contribution of valence bias}\label{the-contribution-of-valence-bias}}

@@

\hypertarget{the-present-study}{%
\subsection{The present study}\label{the-present-study}}

@@

Increased working memory demands (i.e., a higher cognitive load) do not always result in poorer performance on concurrent tasks. For instance Baddeley -({\textbf{???}}) reported that increasing load by adding digits to a rehearsed number did not affect accuracy on a concurrent verbal reasoning task--instead, there was an increase in the latency of response, a potential interference effect that did not alter overall accuracy.

@@

@@
